%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Preámbulo                                                                    %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt,a4paper,titlepage,twoside,openright,openbib]{report}

\usepackage{styletfm}

\usepackage{parskip}


\hypersetup{pdfauthor={Nombre del estudiante},
            pdftitle={Título del trabajo fin de Máster},
            pdfsubject={Traballo Fin de Máster},
            pdfkeywords={Palabrasclave}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Cuerpo                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Preliminares del documento           %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 \include{portada/portada}
 \paginaenblanco
 \dedicatoria{To Mr. Metzger, my mathematics teacher, who helped shape my thinking and analytical skills.}
 \paginaenblanco
 \paginaenblanco
 \begin{Acknowledgements}
 	
I would like to express my sincere gratitude to all my professors throughout my academic journey, and in particular to Juan Angel Lorenzo del Castillo, Karima, and my master's thesis supervisor, Guillermo López Taboada, for their guidance and support.

I would also like to thank Léo De sario, Yvan Moalic, and Cédric Hillion for being by my side during the years in which I built who I am.

I am deeply grateful to Gauthier Dewolf, Yu Tsukimura, Mathys Rey and Tristan Noterman for standing by me through the mentally hardest years of my life.

I am especially thankful to Thomas Brasdefer, Enzo Fernandez, and Antoine Kraus for accompanying me throughout these three years of study.

I would also like to thank Thomas Llasera for encouraging open-mindedness and providing valuable perspective.

A special thank you goes to Clément Cros, who came with me to Spain and shared the experience of the double degree program by my side.

Finally, I am deeply grateful to my parents for always believing in me and for their unwavering support.
	
 \end{Acknowledgements}
 \paginaenblanco
 \include{portada/resumen}
 \paginaenblanco

 \pagenumbering{roman}
 \setcounter{page}{1}

 \tableofcontents
 \listoffigures
 \listoftables
 \cleardoublepage
 
 \pagenumbering{arabic}
 \setcounter{page}{1}

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Capítulos                            %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 \include{contenido/introduccion}
 \include{contenido/StateoftheArt}
 \include{contenido/DatasetandDataPreparation}
 \include{contenido/Methodology}
 \include{contenido/Implementation}
 \include{contenido/ExperimentalResults}
 \include{contenido/Discussion}
 \include{contenido/conclusiones}

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Apéndices, glosarios y bibliografía  %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 \appendix
 \include{anexos/anexos}
%\include{anexos/...}

 \include{bibliografia/acronimos}
 \include{bibliografia/glosario}

 \bibliographystyle{IEEEtranN}
 \bibliography{bibliografia/bibliografia}
 \cleardoublepage
 
\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
